MS797 Glossary

### 2022-02-19 _academics/ms797-glossary/ms797-glossary.Rmd
New terms, knit

variance inflation factor (102)
: "the ratio of the variance of $\hat{\beta_j}$ when fitting the full model divided by the variance of $\hat{\beta_j}$ on its own"; smallest possible value of 1 indicates the absence of collinearity, 5-10 indicates a "problematic amount". Hairy forumla not reproduced.

>>>>

variance inflation factor (102)
: "the ratio of the variance of $\hat{\beta_j}$ when fitting the full model divided by the variance of $\hat{\beta_j}$ on its own"; smallest possible value of 1 indicates the absence of collinearity, 5-10 indicates a "problematic amount". 
$$\text{VIF}(\hat{\beta_j}) = \frac{1}{1 - {R_{X_j|X_-j}^2}}$$
"$R_{X_j|X_-j}^2$ is the $R^2$ from a regression of $X_j$ onto all of the other predictors"

K-nearest neighbors regression (105)
: a mode of regression that seeks to classify observations by their proximity to classified neighbors

curse of dimensionality (107)
: when an observation has no nearby neighbors due to a high number of dimensions exponentially increasing the available space for other observations to be spread out in

# Chapter 4

qualitative / categorical (129)
: interchangeable term for a variable with a non-quantitative value, such as color

classification (129)
: predicting a qualitative response for an observation

classifier (129)
: a classification technique, such as: logistic regression, linear discriminant analysis, quadratic discriminant analysis, naive Bayes, and K-nearest neighbors

logistic function (134)
: a specific function returning an S-shaped curve with values between 0 and 1, used in logistic regression
$$p(X) = \frac{e^{\beta_0 + \beta_1X}}{1 + e^{\beta_0 + \beta_1X}}$$

odds (134)
: the likelihood of a particular outcome: the ratio of the number of results that produce the outcome versus the number that do not, between 0 and $\infty$ (very low or very high probabilities)
$$odds = \frac{p(X)}{1 - p(X)} = e^{\beta_0 + \beta_1X}$$
Note: odds is not the same as probability! Odds of 1/4 means a 20% probability, not 25%.

To convert from odds to probability: divide odds by (1 + odds), as in:

$$\frac{1}{4} \div \left[ 1 + \frac{1}{4} \right] = \frac{1}{4} \div \frac{5}{4} = 1/5 = 0.2$$

To convert from probability to odds, divide probability by (1 - probability), as in:
$$\frac{1}{5} \div \left[ 1 - \frac{1}{5} \right] = \frac{1}{5} \div \frac{4}{5} = \frac{1}{5} \times \frac{5}{4} = 1/4 = 0.25$$
log odds / logit (135)
: the log of the odds 
$$\text{log} \left( \frac{p(X)}{1 - p(X)} \right)$$

likelihood function (135)
: a function that produces the closest possible match of a set of observations, for use in predicting the classifications of other points
hairy equation not rendered

confounding (139)
: when predictors correlate; usually this is something to work to avoid

multinomial logistic regression (140)
: logistic regression into more than 2 categories

softmax coding (141)
: an alternative coding for multiple logistic regression that treats all classes symmetrically, rather than establishing one as a baseline

(probability) density function (142)
: a function returning the likelihood of a particular value occurring in a given sample, between 0 and 1; often used in pairs to establish likelihood of a value within a range rather than the likelihood of an infinitely-thin single "slice"

prior (142)
: the probability that a random observation comes from class $k$

posterior (142)
: the probability that an observation belongs to a class $k$ given its predictor value

normal / Gaussian (143)
: characterized by a bell-shaped curve, not uniformly distributed but tending towards a likeliest/central value

overfitting (148)
: mapping too closely to idiosyncrasies in training data, increasing variance in a model

null classifier (148)
: a classifier that always predicts a zero or null status

confusion matrix (148)
: a matrix showing how many predictions were made, and how accurate the classifications were (predicted x actual, hits on TL-BR diagonal and misses on BL-TR diagonal)

sensitivity (149)
: the percentage of a class (like defaulters) that is correctly identified

specificity (149)
: the percentage of a different class (like non-defaulters) that is correctly identified

ROC curve (150)
: Receiver Operating Characteristics; name of a curve showing the overall performance of a classifier resembling the top left corner of a rounded rectangle; area under the curve (AUC) indicates the percentage of correct classifications; the closer it is to square, the bigger the AUC, the better the classifier

marginal distribution (155)
: the distribution of an individual predictor

joint distribution (155)
: the association between different predictors

kernel density estimator (156)
: "essentially a smoothed version of a histogram"

# Chapter 5

model assessment (197)
: the process of evaluating a model's performance

model selection (197)
: the process of selecting the proper level of flexibility for a model

validation set approach (198)
: randomly dividing a set of observations into a training set and a validation or hold-out set, to assess the test error rate

leave one out cross-validation (LOOCV) (200)
: using a single observation for a validation set and the rest as a training set

k-fold cross-validation (CV) (203)
: dividing a set of observation into $k$ groups of approximately equal size, using the first as a validation set and fitting the method on the remaining $k-1$ folds

bootstrap (209)
: a tool to quantify the uncertainty associated with a given estimator or statistical learning method

sampling with replacement (211)
: picking from a set without removing the picked elements
